---
title: "ESM 204 - Assignment 3"
author: "Jamie Miller & Alex Ivina"
date: "May 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stargazer) # To make nicely formatted table of regression analysis
```

```{r, include = FALSE}
# Read in data
whales <- read_csv("HW3Data.csv")
```


1. Create a linear probability model that predicts a respondent’s probability of voting ‘yes’ on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients.

```{r, include = FALSE}
# Regression - linear probability model
vote_lm <- lm(vote ~ risk + bid + NEP + income + age, data = whales)
vote_lm

# vote = 0.1196977 + 0.0007445(risk) - 0.0010699(bid) + 0.0158639(NEP) + 0.0088282(incomeone_percent) + 0.0027386(incomepoor) + 0.0074891(incomerich) + 0.0467922(incomevery_rich) + 0.0099816(agetofifty) - 0.0201190(agetoforty) - 0.0162261(agetosixty) + 0.0204401(agetothirty)

# (1) “risk”: level of risk reduction, (2) “bid”: annual payment for the household, (3) “vote”: 1 is yes, 0 is no, (4)”NEP” : measure of environmental concern (5)”income”: categorical variable for income level, (6)”age”: categorical variable for age.

summary(vote_lm) # Not sure this is needed, but here it is.


# write formal model equation with table of results.
# - risk, coefficient estimate, SE, p-value, or ***
# round in display


coeff <- vote_lm$coefficients
coeff

vote_lm$coefficients["risk"]

```


```{r, results ='asis'}
# stargazer for the results
table_vote_lm <- stargazer(vote_lm, type = 'html')

# Not sure why it doesn't work; seems like the right code? 
```


2. Based on this regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)

```{r}
# This is a first attempt at question 2 based on the slack thread. I literally have no idea why it needs to be done this way.

# Calculate risk reduction per whale by dividung 20% risk by 5 whales, even though the wording seems to suggest that it is a 20% redutction for a whale. 

risk_per_whale <- 0.20/5 # 20% risk reduction / 5 whales = 4% risk reduction per whale
# multiply each coefficient by 0.04

# Equation
# vote = 0.1196977 + 0.0007445*(risk) - 0.0010699*(bid) + 0.0158639*(NEP) + 0.0088282*(incomeone_percent) + 0.0027386*(incomepoor) + 0.0074891*(incomerich) + 0.0467922*(incomevery_rich) + 0.0099816*(agetofifty) - 0.0201190*(agetoforty) - 0.0162261*(agetosixty) + 0.0204401*(agetothirty)

#prob_vote <- 0.1196977 + 0.0007445*(risk_per_whale) - 0.0010699*(bid) + 0.0158639*(NEP) + 0.0088282*(incomeone_percent) + 0.0027386*(incomepoor) + 0.0074891*(incomerich) + 0.0467922*(incomevery_rich) + 0.0099816*(agetofifty) - 0.0201190*(agetoforty) - 0.0162261*(agetosixty) + 0.0204401*(agetothirty)


# Algebra
# vote - vote + 0.0010699(bid) = 0.1196977 + 0.0007445(risk) - 0.0010699(bid) + 0.0158639(NEP) + 0.0088282(incomeone_percent) + 0.0027386(incomepoor) + 0.0074891(incomerich) + 0.0467922(incomevery_rich) + 0.0099816(agetofifty) - 0.0201190(agetoforty) - 0.0162261(agetosixty) + 0.0204401(agetothirty) - vote + 0.0010699(bid)

# 0.0010699(bid) = 0.1196977 + 0.0007445(risk) - (vote) + 0.0158639(NEP) + 0.0088282(incomeone_percent) + 0.0027386(incomepoor) + 0.0074891(incomerich) + 0.0467922(incomevery_rich) + 0.0099816(agetofifty) - 0.0201190(agetoforty) - 0.0162261(agetosixty) + 0.0204401(agetothirty) 

#prob_bid  <- (0.1196977 + 0.0007445*(risk) - (vote) + 0.0158639*(NEP) + 0.0088282*(incomeone_percent) + 0.0027386*(incomepoor) + 0.0074891*(incomerich) + 0.0467922*(incomevery_rich) + 0.0099816*(agetofifty) - 0.0201190*(agetoforty) - 0.0162261*(agetosixty) + 0.0204401*(agetothirty))/0.0010699
#prob_bid

# calculate for baseline risk X
# calculate for baseline risk X +4%
# turns out they cancel out


### find number and conceptually how you found the number
# how much would bid change with additional 4% risk reduction
# look at code or last HW 

# This equation should do it.
baseline  = -4 * vote_lm$coefficients["risk"]/vote_lm$coefficients["bid"]
baseline

# Every 4% reduction in risk equates to a $2.78 increase in bid

baseline_4  = -4 * vote_lm$coefficients["risk"]/vote_lm$coefficients["bid"] 



```
# Every 4% reduction in risk equates to a $2.78 increase in bid



3. Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR program offering 60% risk reduction.


```{r}

#####################

# Calculate mean vote coefficient
mean_vote <- whales %>% 
  summarize(mean = mean(vote))
mean_vote

# mean values of all votes = 0.714; Sam said to use this value
# Chris said to use 0.5

#####################

# Choose any 3 respondents. Plug in their values to the equation. 

prob_bid_1  <- (0.1196977 + 0.0007445*(60) - (0.714) + 0.0158639*(13) + 0.0088282*(1) + 0.0027386*(0) + 0.0074891*(0) + 0.0467922*(0) + 0.0099816*(1) - 0.0201190*(0) - 0.0162261*(0) + 0.0204401*(0))/0.0010699

#prob_bid_1 = -303.38517618469 <- person that clearly doesn't like whales

prob_bid_2  <- (0.1196977 + 0.0007445*(60) - (0.714) + 0.0158639*(37) + 0.0088282*(0) + 0.0027386*(1) + 0.0074891*(0) + 0.0467922*(0) + 0.0099816*(0) - 0.0201190*(0) - 0.0162261*(0) + 0.0204401*(1))/0.0010699

#prob_bid_2 = 56.5573418076456

prob_bid_3  <- (0.1196977 + 0.0007445*(60) - (0.714) + 0.0158639*(38) + 0.0088282*(0) + 0.0027386*(0) + 0.0074891*(0) + 0.0467922*(1) + 0.0099816*(0) - 0.0201190*(0) - 0.0162261*(1) + 0.0204401*(0))/0.0010699

#prob_bid_3 = 78.2895597719414
```
Respondent 1 is not willing to pay anything for the VSR program since their bid value is negative \$303. Respondent 2 is willing to pay \$56.56 for a VSR program that offers a 60% reduction in whlae strikes, and Respondent 3 is willing to pay \$78.29.



4. Now repeat this process and estimate the mean WTP for a VSR program offering 60% risk reduction among Santa Barbara County households.

```{r}

# Create summy variables for income and age categories
data_new<- (whales) %>% 
  mutate(income_value = case_when(income == "poor" ~ 1,
                                 income == "middle" ~ 2,
                                 income == "rich" ~ 3,
                                 income == "very_rich" ~ 4,
                                 income == "one_percent" ~ 5),
         age_value = case_when(age == "tothirty" ~ 1,
                              age == "toforty" ~ 2,
                              age == "tofifty" ~ 3,
                              age == "tosixty" ~ 4,
                              age == "oversixty" ~ 5)
         )

data_new

# Summarize data. Mean of continuous variables. Median or mode of categorical dummy variables.
data_table <- data_new %>%
  summarize(
    mean_risk = mean(risk),
    median_risk = median(risk),
    mean_bid = mean(bid),
    median_bid = median(bid),
    mean_vote = mean(vote),
    median_vote = median(vote),
    mean_nep = mean(NEP),
    median_nep = median(NEP),
    mean_income = mean(income_value),
    median_income = median(income_value),
    mean_age = mean(age_value),    
    median_age = median(age_value)
  )

data_table 

```

```{r}
# Used pivot table in excel to find mode of all columns for all respondents

# risk = 20
# bid = 20
# vote = 1
# NEP = 40
# income = one_percent
# age = tothirty


```


```{r}
# mean_NEP: 38.366
# most_common_age: agetofifty (I am not 100% sure why it is this)
# most_common_income: medium 

four_mean_714  <- (0.1196977 + 0.0007445*(60) - (data_table$mean_vote) + 0.0158639*(data_table$mean_nep) + 0.0088282*(0) + 0.0027386*(0) + 0.0074891*(0) + 0.0467922*(0) + 0.0099816*(1) - 0.0201190*(0) - 0.0162261*(0) + 0.0204401*(0))/0.0010699
four_mean_714
# WTP: 64.4767617534349

four_mean_5  <- (0.1196977 + 0.0007445*(60) - (0.5) + 0.0158639*(38.366) + 0.0088282*(0) + 0.0027386*(0) + 0.0074891*(0) + 0.0467922*(0) + 0.0099816*(1) - 0.0201190*(0) - 0.0162261*(0) + 0.0204401*(0))/0.0010699
four_mean_5

# WTP: 264.4954.....

# Not exactly sure what is going on here, was just trying to make sense and there is no one at Bren :( 

```


```{r}
# calculared with median NEP, vote = 0.5, risk = 60, mode of income (one_percent) and age(tothirtry)

#bid  <- (0.1196977 + 0.0007445*(risk) - (vote) + 0.0158639*(NEP) + 0.0088282*(incomeone_percent) + 0.0027386*(incomepoor) + 0.0074891*(incomerich) + 0.0467922*(incomevery_rich) + 0.0099816*(agetofifty) - 0.0201190*(agetoforty) - 0.0162261*(agetosixty) + 0.0204401*(agetothirty))/0.0010699


bid_median_5  <- (0.1196977 + 0.0007445*(60) - (0.5) + 0.0158639*(data_table$median_nep) + 0.0088282*(1) + 0.0027386*(0) + 0.0074891*(0) + 0.0467922*(0) + 0.0099816*(0) - 0.0201190*(0) - 0.0162261*(0) + 0.0204401*(1))/0.0010699
bid_median_5

# w/ median $291.92
# w/ mean $282.52

```




```{r}

# Used values from summarized data table. Mean of all continuous variables. Used median of categorical variables. Mode did not work.
# Not 100% certain assigning age_values as actual ages gets us where we want to go. Try again with dummy variables 1-5

mean_bid  <- (0.1196977 + 0.0007445*(60) - (data_table$mean_vote) + 0.0158639*(data_table$mean_nep) + # Continuous variables
                0.0088282*(data_table$median_income) + 0.0027386*(data_table$median_income) + 0.0074891*(data_table$median_income) + 0.0467922*(data_table$median_income) +  # Income categories
                0.0099816*(data_table$median_age) - 0.0201190*(data_table$median_age) - 0.0162261*(data_table$median_age) + 0.0204401*(data_table$median_age)   # Age categories
              )/0.0010699
mean_bid

# mean_bid = -106.2354 when risk = data_table$mean_risk and categorical age variables are as follows: 
#case_when(age == "tothirty" ~ 30,
 #         age == "toforty" ~ 40,
  #        age == "tofifty" ~ 50,
   #       age == "tosixty" ~ 60,
    #      age == "oversixty" ~ 70)
# and income_value =
#case_when(income == "poor" ~ 0,
 #         income == "middle" ~ 1,
  #        income == "rich" ~ 2,
   #       income == "very_rich" ~ 3,
    #      income == "one_percent" ~ 4)


# mean_bid = 161.6301 when risk = 60, income values are the same as above, and categorical age variables are as follows: 
#case_when(age == "tothirty" ~ 1,
 #         age == "toforty" ~ 2,
  #        age == "tofifty" ~ 3,
   #       age == "tosixty" ~ 4,
    #      age == "oversixty" ~ 5)

# Having income$poor = 0 doesn't make sense. Change it to 1. Adjust dummy variables by 1.

# Now mean_bid = 223.1762

# After going through this little exercise I'm not sure this is correct or necessary anyway.
# Answer should be around 50-60 for both Q3 and Q4.

```



5. If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR program to these households.

```{r}

# Risk = 60%  
# Cost = 7 million 
# 150000 * # answer from Q.4

# benefit <- mean_bid * 150000 # Jamie's version; this is wrong. do not use this value

benefit_1 <- four_mean_714 * 150000 # Alex's version; using mean vote = 0.714
benefit_1
# 9,671,514

benefit_2 <- four_mean_5 * 150000 # Alex's version; using median vote = 0.5
benefit_2
# 39,674,318

cost <- 7000000

total_benefit <- benefit_2 - cost
#total_benefit: 32,674,318.2633891
# total benefit: 2,671,514 (does this number seem right or too low? From what people said this maybe should be 11.5?)

# Either way the benefit of the program is greater than the $7 million cost of the VSR program

```

6. Based only on these benefits, do the benefits of the VSR program outweigh the costs?

```{r}

# Depends if the benefit is $2.6 mil or $11.6 -- I think as long as this is a positive number it is ok.


# Yes. Either way the benefit of the program of $9.7 million or $39.7 million is greater than the $7 million cost of the VSR program. The benefits outweigh the costs.

```


7. Suppose the price of a carbon credit is \$Z per ton, at what price Z* will the shipping industry decide to adopt the VSR for purely self-interested reasons? (Note: if ships voluntarily reduce speed, then the \$7 million cost (noted above) of implementing the VSR would be avoided)
```{r}

carbon_credit <- 1000/20 # $1000/20 tons = $50/ton
carbon_credit

# value of whale, shipping cost, carbon credit value
# copmare that with no program and voluntary enrollment


```


8. Now suppose the carbon credit price is exactly Z*, so all ships voluntarily reduce speed to achieve the 60% risk reduction. Approximately how many whales would be saved by this? What would be the social value of allowing ships to enter the carbon trading market?

```{r}



```

